---
title: "Jacek's Wordle Heuristics"
author: "Jacek Rz"
date: "`r Sys.Date()`"
output:
  md_document:
    variant: gfm
knit: (
  function(inputFile, encoding) { 
    rmarkdown::render( 
      input       = inputFile, 
      encoding    = encoding,     
      output_file = 'README')})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(testit)
library(knitr)
```

# Jacek's Wordle Heuristics
Quick analysis of Wordle's dictionary to find out
what are good words to start with and quickly
hit as many target word characters as possible.

```{r words}
words <- read_csv('words.csv', col_names = 'word', col_types = 'c') %>%
  mutate(word = toupper(word))

words.total <- length(words$word)
```

## Character distribution
There are `r words.total` words in the dictionary.
This table shows how many words contain each letter:
```{r chars}
chars = words %>% 
  mutate(char = strsplit(word, '')) %>% 
  unnest_longer(char) %>% 
  unique()

char_freqs <- chars %>% 
  group_by(char) %>%
  summarise(n_words = n()) %>% 
  arrange(desc(n_words))

kable(char_freqs, caption = 'Word count by character')

```


## First guess
```{r first_word}
top_chars <- function(min.inclusive, max.inclusive) {
  char_freqs$char[min.inclusive:max.inclusive]
}

words_from_chars <- function(cs, min.matches = 5) {
  chars %>% 
    filter(char %in% cs) %>% 
    group_by(word) %>% 
    summarise(n = n()) %>% 
    filter(n >= min.matches) %>% 
    .$word
}

coverage_of_chars <- function(cs, min.matches = 5) {
  n <- words_from_chars(cs, min.matches = min.matches) %>% 
    length()
  frac <- n / words.total
  paste(round(frac * 100), '%', sep='')
}
```

Characters **`r top_chars(1, 5)`** are the most frequent ones.
They are all contained by each one of these words:

```
`r words_from_chars(top_chars(1, 5))`
```

With any of the above words as the first guess,
there is a **`r coverage_of_chars(top_chars(1, 5), min.matches = 1)`** chance that at least a single character from
the day's word will be hit, **`r coverage_of_chars(top_chars(1, 5), min.matches = 2)`** chance for unveiling two characters,
and **`r coverage_of_chars(top_chars(1, 5), min.matches = 3)`** chance for hitting as many as three.

## Second guess
Characters **`r top_chars(6, 10)`** come as the next group in terms of frequency.
It's possible to form a 5-character word from all of them too:

```
`r words_from_chars(top_chars(6, 10))`
```

At this point two characters are known in **`r coverage_of_chars(top_chars(1, 10), min.matches = 2)`** of cases.
There is a **`r coverage_of_chars(top_chars(1, 10), min.matches = 3)`** chance that at least three letters
have been hit and **`r coverage_of_chars(top_chars(1, 10), min.matches = 4)`** likelihood of having unveiled
four letters.

## Third guess?!
Still stuck after the first two blind guesses? This is the next group of characters based on frequency:
**`r top_chars(11, 15)`**

It's not possible to form a word from just these characters.
In fact, it's not possible to create a word from any five of the so far unused characters.
So let's find words that introduce four new letters:

```
`r words_from_chars(top_chars(11, 15), min.matches = 4)`
```

After these three guesses, we know at least three characters of the day's word
in **`r coverage_of_chars(top_chars(1, 14), min.matches = 3)`** of games.
That's not bad, is it?